### 2016-08-08 docs/latex/ver1.tex
added generative model doc

\documentclass[12pt]{article}
\pdfoutput=1

\addtolength{\oddsidemargin}{-.875in}
\addtolength{\evensidemargin}{-.875in}
\addtolength{\textwidth}{1.75in}

\addtolength{\topmargin}{-.875in}
\addtolength{\textheight}{1.75in}

\openup 1em

%macro for commenting
\usepackage{color}
\newcommand{\leo}[1]{{\color{blue}{\it leo: #1}}}
\newcommand{\xbeta}{\boldsymbol X_i^T \boldsymbol\beta}

\usepackage[round]{natbib}

\usepackage{rotating}
\usepackage{graphicx}
\usepackage{subcaption}

\usepackage{float}

\usepackage{amsthm,amsmath} 
\usepackage{amssymb}
\usepackage{subcaption}
\captionsetup{compatibility=false}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

\thispagestyle{empty}
\baselineskip=28pt

\title
    {{jk-means}}

\date{}

\begin{document}

\maketitle

\section{Generative Model}

For $i=1\ldots n$, 
\begin{equation}
\begin{aligned}
(w_{1},\ldots,w_{K}) & \sim Dir(\alpha,\ldots,\alpha)\\
Pr(u_{i1},\ldots,u_{iK}) & = \frac{1}{\binom{K}{J}} \\
m_i& \sim Multi(w_1 u_{i1},\ldots,w_K u_{iK})\\
y_i& \sim G(\theta_{m_i})
\end{aligned}
\end{equation}
where $(u_{i1},\ldots,u_{iK})$ is a random sample vector with $J$ $1$'s and $(K-J)$ 0's.

\bibliographystyle{plainnat}
\bibliography{reference}

\end{document}