### 2016-08-29 Rmd/bigKtest.Rmd
First draft of a K=10 test :construction: [neurodata/weekly-experiments#196]

---
title: "$jk$-means tests"
author: "Jesse Leigh Patsolic"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    fig_retina: 2.7
    highlight: pygments
    keep_md: yes
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
---

```{r cc-setup, eval=FALSE, include=FALSE}
require(rmarkdown)
rmarkdown::render("bigKtest.Rmd")
system('open bigKtest.html')
```

```{r knitrOpts, include=FALSE, results='asis', message=FALSE, warning=FALSE}
require(rmarkdown)
require(knitr)
knitr::opts_chunk$set(cache=TRUE, autodep=TRUE)
dep_auto() # figure out dependencies automatically
opts_chunk$set(cache=TRUE, echo=TRUE, 
               opts_chunk$set(fig.path = './'),
               warning=FALSE, message=FALSE, 
               comment="#",
               dpi=227,dev=c('png','pdf'))
```

```{r include=FALSE}
require('ggplot2')
require('reshape')
require(mvtnorm)
require(plotly)
require(jkmeans)
require(mclust)
require(gridExtra)
source("~/neurodata/synaptome-stats/Code/kmpp.r")
```

# Generate Data

Let $X \sim \sum_i^k \pi_i \mathcal{N}_i(\mu_i, \sigma^2_i I)$. 
With $k = 10$ and $n = 1000$, $N = nk$.

```{r cc-genParams}
n <- 1e3
k <- 10
N <- n*k

set.seed(49)
(mux <- sort(sample(-20:20, 10)))
(muy <- sample(-20:20, 10))

(mu0 <- cbind(mux, muy))
(sigma1 <- sample(c(1,1,1,2,3,4,5), 10, replace=TRUE))
sigma0 <- lapply(sigma1, function(x) matrix(x*c(1,0,0,1), 2, 2))

Z <- rep(1:k, each = n)
```

Here are the means
```{r p-means}
plot(mu0, col = 1:10, pch = 19, cex = 3)
```

We now generate the data with 
```{r cc-genXs}
X0 <- data.frame(t(sapply(Z, function(x) rmvnorm(1, mu0[x,], sigma0[[x]]))), Z)
names(X0) <- c('x1', 'x2', 'Z')

head(X0[sample(dim(X0)[1]),])
dat <- as.matrix(X0[, 1:2])
```

And here are the data colored with the truth.
```{r p-X0s}
ggCol <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a')
ggplot(data = X0, aes(x=x1, y=x2, color = as.factor(Z))) +
  scale_color_manual(values = ggCol) + 
  geom_point(alpha = 0.5)
```

# Run algorithms

## Run $jk$-means
Run with $jk$-means with $j=2, k=10$:

options: 
max iteration=100
convergence error tolerance = 1E-8
```{r}
K <- k
j <- 5

set.seed(5431)
set.seed(2^10)
system.time(
  jk <- jkmeans::jkmeansEM(dat, K, j , 100, tol = 1E-8)
)
jkv <- apply(jk$zeta, 1,  function(x) which( x == max(x) ))
```

## Run $k=10$-means initialized at the truth

```{r cc-km}
system.time({
  kv <- kmeans(dat, centers = mu0)
}
)
```

## Run MCLUST

```{r cc-mclust}
system.time(
  mcv <- Mclust(dat, 
              G = K, 
              modelNames="EII")
)
mcv
```
# Results

```{r, echo=FALSE}
ari <- c(mclust::adjustedRandIndex(Z, kv$cluster),
mclust::adjustedRandIndex(Z, jkv),
mclust::adjustedRandIndex(Z, mcv$classification))

Names <- c("Truth v km", "Truth v jk", "Truth v mclust")

data.frame(Names, ari)
```

```{r p-results}
Y <- data.frame(X0, jk=jkv, kv = kv$cluster,
                mc=mcv$classification, kvpp=kvpp$cluster)

p1 <- ggplot(data = Y, aes(x=x1, y=x2)) +
  scale_color_manual(values = ggCol) 
```

```{r p-all, echo=FALSE, fig.height=12,fig.width=16}
layoutm <- matrix(c(NA,2,NA,3,1,4,NA,NA,NA),3,3)

grid.arrange(
 p1 + geom_point(alpha = 0.25, aes(color = as.factor(Z))) +
   ggtitle("Truth"),
 p1 + geom_point(alpha = 0.25, aes(color = as.factor(jkv))) +
   ggtitle("(2,10)=jk-means"),
 p1 + geom_point(alpha = 0.25, aes(color = as.factor(mc))) +
   ggtitle("10=k, mclust"),
 p1 + geom_point(alpha = 0.25, aes(color = as.factor(kv))) +
   ggtitle("10=k-means"),
# p1 + geom_point(alpha = 0.25, aes(color = as.factor(kvpp))) +
#   ggtitle("10=k-meanspp"), 
 layout_matrix = layoutm
)
```