### 2016-09-06 Rmd/convergenceRates.Rmd
sync

```
###Prerequisite
###Install the package if you haven't
setwd("~/git/")
require('devtools')
build('jkmeans')
install.packages("jkmeans_1.0.tar.gz", repos = NULL, type = "source")
```

```{r include=FALSE}
require('ggplot2')
require('reshape')
require('jkmeans')
```

###1. Generate Data
Let's generate data from 2 clusters, with size n/2, n/2*3

N( c(0.7), diag(1.7661/30))
N( c(0.3), diag(1.7661/30))

```{r}
n<- 100
K<- 2
p<- 1
batchN<- 1000

yBatch<- array(0,dim = c(n*K,p,batchN))

mu0<- c(c(0.7), c(0.3))
mu<- matrix( 0,n*2,p)
mu[1:(n/2),]<- rep(c(0.7),each=(n/2))
mu[(n/2+1):(2*n),]<- rep(c(0.3),each=(n/2*3))

sigma<-  sqrt(1.7661/30)

for(b in 1:batchN){
  yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
}

mu_ini<- matrix(mu0,K,p)

```

Here is a view of the two clusters
```{r}

dataHist<- data.frame("Mu"=as.factor(mu),    "y"= yBatch[,,1])

ggplot(dataHist, aes(y, fill = as.factor(Mu))) + geom_histogram(alpha = 0.2,bins = 30,  position="identity")

```

functions to compute misclassification error and rmse of mean estimate:
NB: GMM can get stuck in saddle point at (w1=0, w2=1), so I removed the test result when it happened.
```{r}
computeMError<- function(jk, n,batchN){
  error<- numeric(batchN)
  for(b in 1:batchN){
    newOrder<- order(jk$mu[,1,b],decreasing = T)
    tM0<- c(0,1)[newOrder]
    trueM<- c(rep(tM0[1],(n/2)),rep(tM0[2],(n/2*3)))
    error[b]<- sum (jk$M[,b] != trueM) /2/n
    if( min(jk$w[,b])<0.1){
        error[b]<- NA
    }
  }
  error
}

computeRMSE<- function(jk, mu0, batchN){
  rmse<- numeric(batchN)
  for(b in 1:batchN){
    newOrder<- order(jk$mu[,1,b],decreasing = T)

    muEst<-as.matrix(jk$mu[,,b])[newOrder,]

    rmse[b]<- sqrt( mean((muEst - mu0)^2))
    if( min(jk$w[,b])<0.1)
      rmse[b]<- NA
  }
  rmse
}

```

###2. Test jk-means/ j-sparse-GMM with different n

```{r}

experiment<- function(n){

  K<- 2
  p<- 1
  batchN<- 1000

  yBatch<- array(0,dim = c(n*K,p,batchN))

  mu0<- c(c(0.7), c(0.3))
  mu<- matrix( 0,n*2,p)
  mu[1:(n/2),]<- rep(c(0.7),each=(n/2))
  mu[(n/2+1):(2*n),]<- rep(c(0.3),each=(n/2*3))

  sigma<-  sqrt(1.7661/30)

  for(b in 1:batchN){
    yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
  }

  mu_ini<- matrix(mu0,K,p)

  jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
  jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
  jGMM12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)
  jGMM22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)

  error <- cbind( computeMError(jk12,n, batchN), computeMError(jk22,n,batchN), computeMError(jGMM12,n,batchN),computeMError(jGMM22,n,batchN))
  rmse <- cbind( computeRMSE(jk12,mu0,batchN), computeRMSE(jk22,mu0,batchN), computeRMSE(jGMM12,mu0,batchN),computeRMSE(jGMM22,mu0,batchN))

  list("MCE"=error, "RMSE"=rmse)

}

```

```{r}
#code to run on an increasing series of n

if(FALSE)
{
  nSeries<- c(seq(20,100,by = 20),150,200,300,400,500)

MCEmean<- numeric()
MCEq25<- numeric()
MCEq975<- numeric()
RMSEmean<- numeric()
RMSEq25<- numeric()
RMSEq975<- numeric()

for(n in nSeries){
  test<-experiment(n)

  MCEmean <- rbind(MCEmean,colMeans(test$MCE,na.rm = T))
  MCEq25 <- rbind(MCEq25, apply(test$MCE, MARGIN = 2, function(x){quantile(x,probs = 0.025,na.rm = T)}))
  MCEq975 <- rbind(MCEq975, apply(test$MCE, MARGIN = 2, function(x){quantile(x,probs = 0.975,na.rm = T)}))

  RMSEmean<- rbind(RMSEmean, colMeans(test$RMSE,na.rm = T))
  RMSEq25<- rbind(RMSEq25, apply(test$RMSE, MARGIN = 2, function(x){quantile(x,probs = 0.025,na.rm = T)}))
  RMSEq975<-  rbind(RMSEq975, apply(test$RMSE, MARGIN = 2, function(x){quantile(x,probs = 0.975,na.rm = T)}))

  print(n)
}

result<- list("MCEmean"=MCEmean,
     "MCEq25"=MCEq25,
     "MCEq975"=MCEq975,
     "RMSEmean"=RMSEmean,
     "RMSEq25"=RMSEq25,
     "RMSEq975"=RMSEq975)

save(result, file="resultRates.Rda")
}

```

```{r}

load("resultRates.Rda")

```