### 2016-10-05 Rmd/convergenceRatesImbalancedLargeSigma.Rmd
update large and small overlap

---
title: "jk test with unbalanced label"
author: "Author"
date: '`r Sys.Date()`'
output:
  md_document: 
    variant: markdown_github
---

```
###Prerequisite
###Install the package if you haven't
setwd("~/git/")
require('devtools')
build('jkmeans')
install.packages("jkmeans_1.0.tar.gz", repos = NULL, type = "source")
```

```{r include=FALSE}
require('ggplot2')
require('reshape')
require('jkmeans')

setwd("~/git/jkmeans/Rmd/")
```

###1. Generate Data
Let's generate data from 2 clusters, with size n/2, n/2*3

N( c(0.3), diag(1.7661/30))
N( c(0.7), diag(1.7661/30))

```{r}
n<- 100
K<- 2
p<- 1
batchN<- 1000

yBatch<- array(0,dim = c(n*K,p,batchN))

mu0<- c(c(0.3), c(0.7))
mu<- matrix( 0,n*2,p)
mu[1:(n/2),]<- rep(c(0.3),each=(n/2))
mu[(n/2+1):(2*n),]<- rep(c(0.7),each=(n/2*3))

sigma<-  sqrt(1.7661/10)

for(b in 1:batchN){
  yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
}

mu_ini<- matrix(mu0,K,p)

```

Here is a view of the two clusters
```{r}

dataHist<- data.frame("Mu"=as.factor(mu),    "y"= yBatch[,,1])

ggplot(dataHist, aes(y, fill = as.factor(Mu))) + geom_histogram(alpha = 0.2,bins = 30,  position="identity")

```

functions to compute misclassification error and rmse of mean estimate:
NB: GMM can get stuck in saddle point at (w1=0, w2=1), so I removed the test result when it happened.
```{r}

computeMError<- function(jk, n,batchN){
  error<- numeric(batchN)
  for(b in 1:batchN){
    tM0<- rank(jk$mu[,1,b])-1
    trueM<- c(rep(tM0[1],(n/2)),rep(tM0[2],(n/2*3)))
    error[b]<- sum ( jk$M[,b] != trueM) /K/n
    if( min(jk$w[,b])<0.1){
        error[b]<- NA
    }
  }
  error
}

computeRMSE<- function(jk, mu0, batchN){
  rmse<- numeric(batchN)
  for(b in 1:batchN){
    newOrder<- order(jk$mu[,1,b],decreasing = F)

    muEst<-as.matrix(jk$mu[,,b])[newOrder,]

    rmse[b]<- sqrt( mean((muEst - mu0)^2))
    if( min(jk$w[,b])<0.1)
      rmse[b]<- NA
  }
  rmse
}

```

###2. Test jk-means/ j-sparse-GMM with different n's

```{r}

experiment<- function(n){

  K<- 2
  p<- 1
  batchN<- 1000

  yBatch<- array(0,dim = c(n*K,p,batchN))

  mu0<- c(c(0.3), c(0.7))
  mu<- matrix( 0,n*2,p)
  mu[1:(n/2),]<- rep(c(0.3),each=(n/2))
  mu[(n/2+1):(2*n),]<- rep(c(0.7),each=(n/2*3))

  sigma<-  sqrt(1.7661/10)

  for(b in 1:batchN){
    yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
  }

  mu_ini<- matrix(mu0,K,p)

  jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
  jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
  jGMM12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)
  jGMM22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)

  error <- cbind( computeMError(jk12,n, batchN), computeMError(jk22,n,batchN), computeMError(jGMM12,n,batchN),computeMError(jGMM22,n,batchN))
  rmse <- cbind( computeRMSE(jk12,mu0,batchN), computeRMSE(jk22,mu0,batchN), computeRMSE(jGMM12,mu0,batchN),computeRMSE(jGMM22,mu0,batchN))

  list("MCE"=error, "RMSE"=rmse)

}

```

```{r}
#code to run on an increasing series of n
#ran on cluster

nSeries<- c(seq(20,500,by = 20))

if(FALSE)
{

  MCEmean<- numeric()
  MCEq25<- numeric()
  MCEq975<- numeric()
  RMSEmean<- numeric()
  RMSEq25<- numeric()
  RMSEq975<- numeric()

  for(n in nSeries){
    test<-experiment(n)

    MCEmean <- rbind(MCEmean,colMeans(test$MCE,na.rm = T))
    MCEq25 <- rbind(MCEq25, apply(test$MCE, MARGIN = 2, function(x){quantile(x,probs = 0.025,na.rm = T)}))
    MCEq975 <- rbind(MCEq975, apply(test$MCE, MARGIN = 2, function(x){quantile(x,probs = 0.975,na.rm = T)}))

    RMSEmean<- rbind(RMSEmean, colMeans(test$RMSE,na.rm = T))
    RMSEq25<- rbind(RMSEq25, apply(test$RMSE, MARGIN = 2, function(x){quantile(x,probs = 0.025,na.rm = T)}))
    RMSEq975<-  rbind(RMSEq975, apply(test$RMSE, MARGIN = 2, function(x){quantile(x,probs = 0.975,na.rm = T)}))

    print(n)
  }

  result<- list("MCEmean"=MCEmean,
                "MCEq25"=MCEq25,
                "MCEq975"=MCEq975,
                "RMSEmean"=RMSEmean,
                "RMSEq25"=RMSEq25,
                "RMSEq975"=RMSEq975)

  save(result, file="resultRatesImbalancedLargeSigma.Rda")
}

```

#Misclassification error
plot without & with pointwise 95% confidence band

```{r}

load("resultRatesImbalancedLargeSigma.Rda")

plot1<- data.frame( "n" = rep(nSeries, 4),
            "MC14Error"= c(result$MCEmean),
            "MC14ErrorL"= c(result$MCEq25),
            "MC14ErrorU"= c(result$MCEq975),
            "Model"= rep(c("K-Means","W-constraint-GMM (22-Means)","1-sparse-GMM","GMM"),each= length(nSeries))
            )

p<- ggplot(data=plot1, aes())
p<- p+  geom_line(aes(x= n, y=MC14Error, colour=Model)) 
pE<- p+  geom_line(aes(x= n, y=MC14ErrorL, colour=Model),linetype=2) 
pE<- pE+  geom_line(aes(x= n, y=MC14ErrorU, colour=Model),linetype=2)

# p+ geom_errorbar( aes(x=n,ymax = MC14ErrorU, ymin=MC14ErrorL, colour=Model),width=0.2)
p+ theme_bw()
pE + theme_bw()
```

#RMSE for the mean estimate
plot without & with pointwise 95% confidence band

```{r}

plot1<- data.frame( "n" = rep(nSeries, 4),
            "RMSE"= c(result$RMSEmean),
            "RMSEL"= c(result$RMSEq25),
            "RMSEU"= c(result$RMSEq975),
            "Model"= rep(c("K-Means","W-constraint-GMM (22-Means)","1-sparse-GMM","GMM"),each= length(nSeries))
            )

p<- ggplot(data=plot1, aes())
p<- p+  geom_line(aes(x= n, y=RMSE, colour=Model))

pE<- p+  geom_line(aes(x= n, y=RMSEL, colour=Model),linetype=2) 
pE<- pE+  geom_line(aes(x= n, y=RMSEU, colour=Model),linetype=2)

p+ theme_bw()
pE + theme_bw()

```