### 2016-11-07 Rmd/HighP.Rmd
updated write-up, test with large p

# .rs.restartR()
```{r include=FALSE}

require("jkmeans")

n<- 100
K<- 2
p<- 100

sigma<- 1
mu1<- rep(K,n*K)

randomMu<- unlist(sapply(1:K, function(x){rep(x, runif(1,n/10,n))}))
mu1[1:length(randomMu)]<- randomMu

mu<- matrix(0,n*K,p)
mu[,1:5]<- mu1
# mu[,2]<- mu1

y<- matrix( rnorm(n*K*p, mu,sd = (sigma)), n*K)
# hist(y,breaks = 100)

svdY<- svd(y)

plot(svdY$d)

y1<- as.matrix( t(t(svdY$u)*svdY$d) [,c(1)])

# hist(c(sapply(c(1:K),function(x){ rnorm(1000,x, (sigma2/sqrt(n)))})), breaks = 1000)

kmpp <- function(X, k) {
  n <- nrow(X)
  C <- numeric(k)
  C[1] <- sample(1:n, 1)

  for (i in 2:k) {
    dm <- abs(outer(X, X[C, ], "-"))
    pr <- apply(dm, 1, min)
    pr[C] <- 0
    C[i] <- sample(1:n, 1, prob = pr)
  }

   X[C, ]
}

rmse<- function(x){
  v<- x$mu[,1:5]
  v<- v[order(v[,1]),]
  sqrt(mean((v-c(1:2))^2))
}

plot(y,col=mu1)

k<- K

ini<- as.matrix(kmpp(y,k))

kmeans <- jkmeansEM(y,k = k,j= 1,1000,tol = 1E-15,fixW = T, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
# table(kmeans$M)
# lines(kmeans$mu, type="p",pch=2,lwd=3,col="green")

jk<- jkmeansEM(y,k = k,j= 1,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
# table(jk$M)
# lines(jk$mu, type="p",pch=2,lwd=3,col="blue")
jk$mu[,1]

full<- jkmeansEM(y,k = k,j= k,1000,tol = 1E-10,fixW = F, meansIni = ini,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
# table(full$M)
# lines(full$mu, type="p",pch=3,lwd=3,col="red")
full$mu[,1]

y<- y1
ini<- as.matrix(kmpp(y,k))

kmeans_pca <- jkmeansEM(y,k = k,j= 1,1000,tol = 1E-15,fixW = T, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
# table(kmeans$M)
# lines(kmeans$mu, type="p",pch=2,lwd=3,col="green")
# kmeans$mu[,1]

jk_pca<- jkmeansEM(y,k = k,j= 1,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
# table(jk$M)
# lines(jk$mu, type="p",pch=2,lwd=3,col="blue")
# jk$mu[,1]

full_pca<- jkmeansEM(y,k = k,j= k,1000,tol = 1E-10,fixW = F, meansIni = ini,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
# table(full$M)
# lines(full$mu, type="p",pch=3,lwd=3,col="red")
# full$mu[,1]

((kmeans$mu[1,1]*svdY$v)[,1]-(kmeans$mu[2,1]*svdY$v)[,1])[1:5]
((jk$mu[1,1]*svdY$v)[,1]-(jk$mu[2,1]*svdY$v)[,1])[1:5]
((full$mu[1,1]*svdY$v)[,1] - (full$mu[2,1]*svdY$v)[,1])[1:5]

rmse_pca<- function(x){
  v<-rbind( (x$mu[1,1]*svdY$v)[1:5,1],  (x$mu[2,1]*svdY$v)[1:5,1])
  v<- v[order(v[,1]),]

  sqrt(mean((v-c(1:2))^2))
}

kmeans_pca$mu 
dim(svdY$v)

```

```{r}
rmse(kmeans)
rmse(jk)
rmse(full)
rmse_pca(kmeans_pca)
rmse_pca(jk_pca)
rmse_pca(full_pca)
```