### 2017-01-02 test/eigenface2.Rmd

```{r}
require("pixmap")
require("jkmeans")
require("mclust")

setwd("~/git/jkmeans/lfwcrop_face/lfwcrop_grey/pick/")

faces<- sapply(list.files("."),function(f){
  x=read.pnm(file = f)
  c(x@grey)
})

mFaces<- 0# rowMeans(faces)
faces<- faces-mFaces

M<- sapply(list.files("."), function(x){substr(x,1,5)})
faces<- faces[,M=="Alvar"]

X<- t(faces)

nf<- 64

K<- 4

plotFace<- function(x){
  image(t(matrix(rev(x),64,64)),col = grey(seq(0, 1, length = 256)))
}

raw<- jkmeansEM(X,k = K,useKmeansIni = T,meansIni = matrix(rep(1,2*10),nrow = 2),sigma2_ini = 0.1,fixW = F)

plotFace(raw$mu[1,] + mFaces)
plotFace(raw$mu[2,]+ mFaces)
plotFace(raw$mu[3,]+ mFaces)
plotFace(raw$mu[4,]+ mFaces)
# plotFace(raw$mu[5,]+ mFaces)
# plotFace(raw$mu[6,]+ mFaces)
# plotFace(raw$mu[7,]+ mFaces)
# plotFace(raw$mu[8,]+ mFaces)

plot(raw$M,title="classification")

arcM<- rDARC(X,d = 9,k = K,meansIni = matrix(1,2),steps = 2E3,sigma2_ini = 1,randomStart = F,ver = 2,useEstep = F,tol = 1E-9)

plotFace(arcM$mu[1,]%*%arcM$EV+ mFaces)
plotFace(arcM$mu[2,]%*%arcM$EV+ mFaces)
plotFace(arcM$mu[3,]%*%arcM$EV+ mFaces)
plotFace(arcM$mu[4,]%*%arcM$EV+ mFaces)

plot(arcM$M,title="classification")
```