### 2017-01-09 test/mnist_test.Rmd
tested GMM/kmeans/1-max on eigenface mnist

---
title: "GMM, ECM and kmeans"
author: "Author"
date: '`r Sys.Date()`'
output:
  md_document: 
    variant: markdown_github
---

```{r}
# .rs.restartR()
setwd("~/git/jkmeans/test/")
source("kmpp.r")
require("jkmeans")
require("mclust")

mnist<- read.csv("../mnist_data/mnist.csv",header = F)

pick<- c(0:5)
p<- 28^2

n<-100
s<- sample(1:1000,n,replace = F)

subset<- c(sapply(pick*1000, function(x){x+ s}))
mnist_subset<- mnist[,subset]
Y<-t(mnist_subset)
Y<- Y/255
K<- length(pick)

gmm<- jkmeansEM(Y,k = K,useKmeansIni = T,meansIni = matrix(rnorm(K*p),nrow = K),sigma2_ini = 0.1,fixW = F,tau = 1,steps = 1000)

jk<- jkmeansEM(Y,k = K,useKmeansIni = T,meansIni = matrix(rep(1,K*p),nrow = K),sigma2_ini = 0.1,fixW = F,tau = 1000)

kmeans <- jkmeansEM(Y,k = K,useKmeansIni = T,meansIni = matrix(rep(1,K*p),nrow = K),sigma2_ini = 0.1,fixW = T,tau = 1000)

M<- rep(pick,each=n)

adjustedRandIndex(jk$M,M)
adjustedRandIndex(gmm$M,M)
adjustedRandIndex(kmeans$M,M)

plot(jk$M)
```